MCC Models

### 2019-10-04 papers/full_paper/manuscript.Rmd
update manuscript

---
title             : "Pragmatic cue integration in adults’ and children’s inferences about novel word meanings"
shorttitle        : "Pragmatic cue integration"

@@

abstract: |
  Language is learned in complex social settings where listeners must reconstruct speakers’ intentend meanings from context. To navigate this challenge, children can use contextual or pragmatic reasoning to learn the meaning of unfamiliar words. One important challenge for pragmatic reasoning is that it often requires integrating information about the current utterance with broader common ground built up over the course of an interaction. Here we study this integration process. We isolate these two sources of pragmatic information and formalize both how they should be combined and how they might develop using a probabilistic model of conversational reasoning. We present a series of seven experiments with three- to five-year-old children and adults that suggest that these two information sources flexibly trade off with one another. A cue combination model accurately predicted empirical results for children and adults and provided a better fit compared with alternative models in which only one information source was considered. This work integrates distinct sets of findings regarding early language and suggests that pragmatic reasoning models can provide a quantitative framework for understanding developmental changes in language learning.

keywords          : "language acquisition, social cognition, pragmatics, Bayesian modeling"
wordcount         : "X"

@@

csl               : pnas.csl

@@

# Significance statement

In order to learn the meaning of a new word, children need to integrate multiple sources of information. Some information is provided by the utterance itself, but other contextual information accumulates over the course of a conversation and is stored in common ground. To understand how these sources are integrated with one another and how they might develop, we formalize this process using computational models of pragmatic reasoning. These models make quantitative predictions for different hypotheses about how integration proceeds. We find that the way in which preschoolers and adults integrate information is best described by a model that flexibly trades off between different sources of information. Taken together, this work presents an explicit theory of socially grounded language learning.

# Introduction

What someone means by an utterance is oftentimes not reducible to the words they used. It takes  pragmatic inference – context-sensitive reasoning about the speaker’s intentions - to recover the intended meaning [@grice1991studies, @levinson2000presumptive, @sperber2001relevance]. Contextual information comes in many forms. On the one hand, there is information provided by the utterance\footnote{We use the terms utterance, utterance-level information or utterance-level cues to capture all cues that the speaker provides for their intended meaning. This includes direct referential information in the form of actions such as pointing or gazing, semantic information in the form of conventional word meanings as well as pragmatic inferences that are licenced by the particular choice of words or actions.} itself. Competent language users expect each other to communicate in a cooperative way such that speakers produce utterances that are relevant and informative. Thus, semantic ambiguity can be resolved by reasoning about why the speaker produced this particular utterance [@grice1991studies, @clark1996using, @sperber2001relevance, @tomasello2008origins]. On the other hand, there is information provided by common ground (the body of knowledge and beliefs shared between interlocutors; [@bohn2018common, @clark1996using, @clark2015common]. Because utterances are embedded in a broader (ongoing) social interaction, common ground information serves as a background against which new utterances are interpreted, making some meanings for ambiguous terms more likely than others.

Children learning their first language make inferences about intended meanings based on utterance-level and common-ground information both for language understanding and language learning [@bohn2019pervasive, @clark2009first, @tomasello2008origins]. Starting very early, infants expect adults to produce utterances in a cooperative way [@behne2005one], and expect language to be carrying information [@vouloumanos2012twelve]. By age two, children are sensitive to the informativeness of communication [@o2001two]. By age three children can use this expectation to make pragmatic inferences [@stiller2015ad, @yoon2019role] and to infer novel word meanings [@frank2014inferring]. And although older children continue to struggle with some complex pragmatic inferences until age five and beyond [@noveck2001children], an emerging consensus identifies these difficulties as stemming from difficulties reasoning about linguistic alternatives rather than pragmatic deficits [@skordos2016children, @horowitz2018trouble, @barner2011accessing]. Thus, children’s ability to reason about utterance-level pragmatics is present at least by ages three to five, and possibly substantially younger.

Evidence for the use of common ground information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social, @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role, @diesendruck2004two, @saylor2009preschoolers, @sullivan2019discourse].

All of these examples, however, highlight children’s use of a single pragmatic information source or cue. Harnessing multiple – potentially competing – cues poses a separate challenge. One aspect of this integration problem is how to balance common ground information that is built up over the course of an interaction against information gleaned from the current utterance. Much less is known about whether and how children – or even adults – combine these types of information. While many theories of pragmatic reasoning presuppose that both information sources are integrated, the nature of their relationship has typically not been specified. We address this challenge here.

Recent innovations in probabilistic models of pragmatic reasoning provide a quantitative method for addressing the problem of integrating multiple sources of contextual information. This class of computational models, which are referred to as Rational Speech Act (RSA) models [@frank2012predicting, @goodman2016pragmatic] formalize the problem of language understanding as a special case of Bayesian social reasoning. A listener interprets an utterance by assuming it was produced by a cooperative speaker who had the goal to be informative. Being informative is defined as providing a message that would increase the probability of the listener recovering the speaker’s intended meaning in context. This notion of contextual informativeness captures the Gricean idea of cooperation between speaker and listener, and provides a first approximation to what we have described above as utterance-level pragmatic information.

Listeners and speakers also enter into a conversation with assumptions about what is likely to be talked about, a reflection of the common ground shared between them. RSA models capture common ground information as a shared prior distribution over possible intended meanings. Thus, a natural locus for information integration within probabilistic models of pragmatic reasoning is the trade off between the prior probability of a meaning and the informativeness of the utterance. This trade off between contextual factors during word learning is a unique aspect of the word learning problem that is not addressed by other computational models of word learning, which have focused on learning from cross-situational, co-occurrence statistics [@fazly2010probabilistic, @frank2009using] or describing generalizations about word meaning [@xu2007word].

We make use of this framework to study pragmatic cue integration across development. To this end, we adapt a method used in perceptual cue integration studies [@ernst2002humans] predictions about conditions in which they either coincide or conflict.

@@

After successfully validating this approach with adults, we apply the same model-driven experimental procedure to children: We first show that they make pragmatic inferences based on utterance and common ground information separately (Experiment 5 and 6). Then we generate a priori model predictions and compare them to data from an experiment – parallel to Experiment 3 – in which both information sources have to be integrated (Experiment 7). To our knowledge, the strategy of independently estimating model parameters with separate groups and combining them via the model to make preregistered quantitative predictions about a new population has not been used before with children.

@@

# How do adults integrate contextual sources of information?

## Inferences based on utterance and common ground information  (Experiments 1 and 2)

@@

In Experiment 1, participants could learn which object a novel word referred to by assuming that the speaker communicated in an informative way [@frank2014inferring]. The speaker was located between two tables, one with two novel objects, A and B, and the other with only object A (Fig. 1A). When the speaker turned and pointed to the table with the two objects (A and B) and used a novel word to request one of them, participants could infer that the word referred to object B. This follows from the counter-factual inferences that, if the (informative) speaker had wanted to refer to object A, they would have pointed to the table with the single object (this being the least ambiguous way to refer to that object). In the control condition, both tables contained both objects and no inference could be made based on the speaker’s behavior. Participants selected object B above chance in the test condition (t(`r ex1t%>%filter(condition == "test")%>%pull(df)`) = `r ex1t%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex1t%>%filter(condition == "test")%>%pull(p_value)`) and more often compared to the control condition (*$\beta$* = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`). 
```{r ex2-results}
ex2tpref <- adult_ex2_preference_data %>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

@@

In Experiments 2A and 2B, we tested if participants use common ground information that is specific to a speaker to identify the referent of a novel word [@akhtar1996role, @saylor2009preschoolers]. In Experiment 2A, the speaker expressed a preference for one of two objects (Fig. 1B, left). Later, the speaker used a novel word to request an object. Adults selected the preferred object above chance (t(`r ex2tpref%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(p_value)`) and more so than in a control condition, where a different speaker, whose preferences were unknown, made the request ($\beta$ = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). In Experiment 2B, common ground information came in the form of novelty (Fig 1B, right). First, the speaker encountered one object on one of the tables. Later, a second object appeared. When the same speaker then used a novel word to request an object, participants selected the new object above chance (t(`r ex2tnov%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(p_value)`), and more often compared to when a different speaker (to whom both objects were equally new) made the request ($\beta$ = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). Taken together, Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model.

## Model predictions for information integration evaluated against new data (Experiment 3).

@@

In the above equation, the listener is trying to jointly resolve the speaker’s intended referent r and the meaning of words (thus learning the lexicon $\mathcal{L}$). He does this by imagining what a rational speaker would say, given the referent they are trying to communicate and a lexicon. The speaker is an approximately rational Bayesian actor (with degree of rationality alpha), who produces utterances as a function of their informativity. The space of utterances the speaker could produce depends upon the lexicon $P(u|\mathcal{L})$; simply put, the speaker labels objects with the true labels under a given lexicon L (see SI Appendix for details):

@@

The speaker’s possible utterances are pairs of linguistic and non-linguistic signals, namely labels and points. Because the listener does not know the lexicon, the informativity of an utterance comes from the speaker’s point, the meaning of which is encoded in $\mathcal{L}_{point}$ and is simply a truth-function checking whether or not the referent is at the location picked out by the speaker’s point. Though the speaker makes their communicative decision assuming the listener does not know the meaning of the labels, we assume that in addition to a point, the speaker produces a label consistent with their own lexicon $\mathcal{L}$, described by $P(u|\mathcal{L})$ (see SI Appendix for modeling details).

This computational model provides a natural avenue to formalize quantitatively how informativeness and common ground trade-off during word learning. As mentioned above, the common ground shared between speaker and listener plays the role of the listener’s prior distribution over meanings, or types of referents, that the speaker might be referring to and which we posit depends on prior interactions around the referents in the present context (e.g., preference or novelty; Experiment 2A and B). We use the results from Experiment 2 to specify this distribution. The in-the-moment, contextual informativeness of the utterance is captured in the likelihood term, whose value depends on the rationality parameter $\alpha$. Assumptions about rationality may change depending on context and we therefore used the data from Experiment 1 to specify $\alpha$ (see SI Appendix for details about these parameters).

The model generates predictions for situations in which utterance and common ground expectations are jointly manipulated (Fig. 1C - see SI Appendix for additional details and a worked example of how predictions were generated). In addition to the parameters fit to the data from previous experiments, we include an additional noise parameter to account for responses better explained by a process of random guessing than by pragmatics; we estimate this parameter from the observed data (Experiment 3). Including the noise parameter greatly improved the model fit to the data (see SI Appendix for details). We did not pre-register the inclusion of a noise parameter for Experiment 3 but did so for all subsequent experiments.
```{r ex3 GLMM, cache=F}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

In Experiment 3, we combined the procedures of Experiment 1 and 2A or 2B. The test setup was identical to Experiment 1, however, before making a request, the speaker interacted with the objects so that some of them were preferred by or new to them (Fig. 1C). We discuss and visualize the results as the proportion with which participants chose the more informative object (i.e., the object that would be the more informative referent when only utterance information is considered). Participants distinguished between congruent and incongruent trials when the speaker remained the same, as evidenced by the fit of a generalized linear mixed effects model (model term: `alignment x speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(p.value)`).

Participants’ average responses were highly correlated with the model’s predictions in each condition (Fig. 2A). To test whether participants in fact balanced both information sources, we compared the pragmatics model to two alternative models: the *flat prior model*, which ignores common ground information and the *prior only model*, which ignores utterance information. Model fit was considerably better for the pragmatics model compared to the flat prior model (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information. The estimated proportion of random responses according to the pragmatics model was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This value was substantially lower for the pragmatics model compared to the alternative models (see SI Appendix), lending additional support to the conclusion that the pragmatics model better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

```{r fig2, include = T, fig.align = "center", fig.cap = "Correlation between model predictions and data. (A) Experiment 3, (B) Experiment 4, (C) 3-year-olds in Experiment 7 and (D) 4-year-olds in Experiment 7. Coefficients and p-values are based on Pearson correlation statistics. Error bars represent  95\\% HDIs.", out.width="450px"}
knitr::include_graphics("./figures/fig2.png")
```

## Replication and extension to different levels of common ground information (Experiment 4).

@@

To test if our model makes accurate predictions for different combinations, we first replicated and then extended the results of Experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (strong, medium and weak manipulation) by changing  the way the speaker interacted with the objects prior to the request We ran a total of 20 conditions, including a direct replication of Experiment 3 (see SI Appendix for details).

Model predictions from the pragmatics model were again highly correlated with the average response in each condition (Fig. 2B). We evaluated model fit for the same models as in Experiment 3 and found again that the pragmatics model fit the data much better compared to the flat prior (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the pragmatics model was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to Experiment 3 and again lower compared to the alternative models (see SI Appendix).

# Do children integrate contextual information?

@@

## Inferences based on utterance and common ground information (Experiment 5 and 6)

The study for children followed the same general pattern as the one for adults. We generated model predictions for how information should be integrated by first measuring children’s ability to use utterance (informativeness) and common ground (preference) information in isolation when making pragmatic inferences. We then adapted our model to study developmental change: We sampled children continuously between 3.0 and 5.0 years of age – a time in which children have been found to make the kind of pragmatic inferences we studied here [@bohn2019pervasive, frank2014inferring] - and generated model predictions for the average developmental trajectory in each condition\footnote{For Experiment 5 and 6, we also tested two-year-olds but did not find sufficient evidence that they use utterance and/or common ground information in the tasks we used to justify investigating their ability to integrate the two.}.

@@

All experiments were framed as games in which participants would learn words from animals. They were implemented in HTML/JavaScript as a website. Adults were directed to the website via MTurk and responded by clicking objects. Children were guided through the game by an experimenter and responded by touching objects on the screen of an iPad tablet [@frank2016using]. For each animal character, we recorded a set of utterances (one native English speaker per animal) that were used to provide information and make requests. All experiments started with an introduction to the animals and two training trials. Subsequent test trials in each condition were presented in a random order. Detailed experimental procedures for each experiment can be found in the SI Appendix.

>>>>

---
title             : "Predicting pragmatic cue integration in adults’ and children’s inferences about novel word meanings"
shorttitle        : "Pragmatic cue integration"

@@

abstract: |
  Language is learned in complex social settings where listeners must reconstruct speakers’ intentend meanings from context. To navigate this challenge, children can use pragmatic reasoning to learn the meaning of unfamiliar words. One important challenge for pragmatic reasoning is that it requires integrating multiple information sources. Here we study this integration process. We isolate two sources of pragmatic information and, using a probabilistic model of conversational reasoning, formalize both how they should be combined and how this process might develop. We use this model to generate quantitative predictions, which we test against new behavioral data from three- to five-year-old children and adults in a series of pre-registered experiments. Results show close numerical alignment between model predictions and data. This work integrates distinct sets of findings regarding early language and suggests that pragmatic reasoning models can provide a quantitative framework for understanding developmental changes in language learning.

keywords          : "language acquisition, social cognition, pragmatics, Bayesian modeling"

@@

csl               : nature.csl

@@

What someone means by an utterance is oftentimes not reducible to the words they used. It takes  pragmatic inference – context-sensitive reasoning about the speaker’s intentions - to recover the intended meaning [@grice1991studies; @levinson2000presumptive; @sperber2001relevance]. Contextual information comes in many forms. On the one hand, there is information provided by the utterance\footnote{We use the terms utterance, utterance-level information or utterance-level cues to capture all cues that the speaker provides for their intended meaning. This includes direct referential information in the form of actions such as pointing or gazing, semantic information in the form of conventional word meanings as well as pragmatic inferences that are licenced by the particular choice of words or actions.} itself. Competent language users expect each other to communicate in a cooperative way such that speakers produce utterances that are relevant and informative. Thus, semantic ambiguity can be resolved by reasoning about why the speaker produced this particular utterance [@grice1991studies; @clark1996using; @sperber2001relevance; @tomasello2008origins]. On the other hand, there is information provided by common ground (the body of knowledge and beliefs shared between interlocutors)[@bohn2018common; @clark1996using; @clark2015common]. Because utterances are embedded in common ground, pragmatic reasoning in context always requires information integration. But how does integration proceed? And how does it develop? Verbal theories assume that information is integrated and that this process develops but do not specify how. We bridge this gap by formalizing information integration and development in a probabilistic model of pragmatic reasoning.

Children learning their first language make inferences about intended meanings based on utterance-level and common-ground information both for language understanding and language learning [@bohn2019pervasive; @clark2009first; @tomasello2008origins]. Starting very early, infants expect adults to produce utterances in a cooperative way [@behne2005one], and expect language to be carrying information [@vouloumanos2012twelve]. By age two, children are sensitive to the informativeness of communication [@o2001two]. By age three children can use this expectation to make pragmatic inferences [@stiller2015ad; @yoon2019role] and to infer novel word meanings [@frank2014inferring]. And although older children continue to struggle with some complex pragmatic inferences until age five and beyond [@noveck2001children], an emerging consensus identifies these difficulties as stemming from difficulties reasoning about linguistic alternatives rather than pragmatic deficits [@skordos2016children; @horowitz2018trouble; @barner2011accessing]. Thus, children’s ability to reason about utterance-level pragmatics is present at least by ages three to five, and possibly substantially younger.

Evidence for the use of common ground information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social; @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role; @diesendruck2004two; @saylor2009preschoolers; @sullivan2019discourse].

All of these examples, however, highlight children’s use of a single pragmatic information source or cue. Harnessing multiple – potentially competing – cues poses a separate challenge. One aspect of this integration problem is how to balance common ground information that is built up over the course of an interaction against information gleaned from the current utterance. Much less is known about whether and how children – or even adults – combine these types of information. While many theories of pragmatic reasoning presuppose that both information sources are integrated, the nature of their relationship has typically not been specified.

Recent innovations in probabilistic models of pragmatic reasoning provide a quantitative method for addressing the problem of integrating multiple sources of contextual information. This class of computational models, which are referred to as Rational Speech Act (RSA) models [@frank2012predicting; @goodman2016pragmatic] formalize the problem of language understanding as a special case of Bayesian social reasoning. A listener interprets an utterance by assuming it was produced by a cooperative speaker who had the goal to be informative. Being informative is defined as providing a message that would increase the probability of the listener recovering the speaker’s intended meaning in context. This notion of contextual informativeness captures the Gricean idea of cooperation between speaker and listener, and provides a first approximation to what we have described above as utterance-level pragmatic information.

Listeners and speakers also enter into a conversation with assumptions about what is likely to be talked about, a reflection of the common ground shared between them. RSA models capture common ground information as a shared prior distribution over possible intended meanings. Thus, a natural locus for information integration within probabilistic models of pragmatic reasoning is the trade off between the prior probability of a meaning and the informativeness of the utterance. This trade off between contextual factors during word learning is a unique aspect of the word learning problem that is not addressed by other computational models of word learning, which have focused on learning from cross-situational, co-occurrence statistics [@fazly2010probabilistic; @frank2009using] or describing generalizations about word meaning [@xu2007word].

We make use of this framework to study pragmatic cue integration across development. To this end, we adapt a method used in perceptual cue integration studies [@ernst2002humans] predictions about conditions in which they either coincide or conflict. Finally, we pre-register these quantitative predictions and test them against new data from adults and children.

@@

After successfully validating this approach with adults, we apply the same model-driven experimental procedure to children: We first show that they make pragmatic inferences based on utterance and common ground information separately (Experiment 5 and 6). Then we generate a priori model predictions and compare them to data from an experiment – parallel to Experiment 3 – in which both information sources have to be integrated (Experiment 7).

@@

# Results

## How do adults integrate contextual sources of information?

### Inferences based on utterance and common ground information  (Experiments 1 and 2)

@@

In Experiment 1, participants could learn which object a novel word referred to by assuming that the speaker communicated in an informative way [@frank2014inferring]. The speaker was located between two tables, one with two novel objects, A and B, and the other with only object A (Fig \ref{fig:fig1}A). When the speaker turned and pointed to the table with the two objects (A and B) and used a novel word to request one of them, participants could infer that the word referred to object B. This follows from the counter-factual inferences that, if the (informative) speaker had wanted to refer to object A, they would have pointed to the table with the single object (this being the least ambiguous way to refer to that object). In the control condition, both tables contained both objects and no inference could be made based on the speaker’s behavior. Participants selected object B above chance in the test condition (t(`r ex1t%>%filter(condition == "test")%>%pull(df)`) = `r ex1t%>%filter(condition == "test")%>%pull(t_value)`, *p* `r ex1t%>%filter(condition == "test")%>%pull(p_value)`) and more often compared to the control condition (*$\beta$* = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`, see Fig \ref{fig:fig2}).

```{r ex2-results}
ex2tpref <- adult_ex2_preference_data %>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

@@

In Experiments 2A and 2B, we tested if participants use common ground information that is specific to a speaker to identify the referent of a novel word [@akhtar1996role; @saylor2009preschoolers]. In Experiment 2A, the speaker expressed a preference for one of two objects (Fig \ref{fig:fig1}B, left). Later, the speaker used a novel word to request an object. Adults selected the preferred object above chance (t(`r ex2tpref%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tpref%>%filter(condition == "same_speaker")%>%pull(p_value)`) and more so than in a control condition, where a different speaker, whose preferences were unknown, made the request ($\beta$ = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). In Experiment 2B, common ground information came in the form of novelty (Fig \ref{fig:fig1}B, right). First, the speaker encountered one object on one of the tables. Later, a second object appeared. When the same speaker then used a novel word to request an object, participants selected the new object above chance (t(`r ex2tnov%>%filter(condition == "same_speaker")%>%pull(df)`) = `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(t_value)`, *p* `r ex2tnov%>%filter(condition == "same_speaker")%>%pull(p_value)`), and more often compared to when a different speaker (to whom both objects were equally new) made the request ($\beta$ = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`, see Fig \ref{fig:fig2}). Taken together, Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model.

```{r fig2, include = T, fig.align = "center", fig.cap = "Results from Experiments 1, 2A, and 2B for adults. For preference and novelty, control refers to a different speaker (see Fig 1B). Transparent dots show data from individual participants, diamonds represent condition means, error bars are 95\\% CIs. Dashed line indicates performance expected by chance.", out.width="300px"}
knitr::include_graphics("./figures/fig2.png")
```

### Model predictions for information integration evaluated against new data (Experiment 3)

@@

In the above equation, the listener is trying to jointly resolve the speaker’s intended referent r and the meaning of words (thus learning the lexicon $\mathcal{L}$). He does this by imagining what a rational speaker would say, given the referent they are trying to communicate and a lexicon. The speaker is an approximately rational Bayesian actor (with degree of rationality alpha), who produces utterances as a function of their informativity. The space of utterances the speaker could produce depends upon the lexicon $P(u|\mathcal{L})$; simply put, the speaker labels objects with the true labels under a given lexicon L (see supplementary information for details):

@@

The speaker’s possible utterances are pairs of linguistic and non-linguistic signals, namely labels and points. Because the listener does not know the lexicon, the informativity of an utterance comes from the speaker’s point, the meaning of which is encoded in $\mathcal{L}_{point}$ and is simply a truth-function checking whether or not the referent is at the location picked out by the speaker’s point. Though the speaker makes their communicative decision assuming the listener does not know the meaning of the labels, we assume that in addition to a point, the speaker produces a label consistent with their own lexicon $\mathcal{L}$, described by $P(u|\mathcal{L})$ (see supplementary information for modeling details).

This computational model provides a natural avenue to formalize quantitatively how informativeness and common ground trade-off during word learning. As mentioned above, the common ground shared between speaker and listener plays the role of the listener’s prior distribution over meanings, or types of referents, that the speaker might be referring to and which we posit depends on prior interactions around the referents in the present context (e.g., preference or novelty; Experiment 2A and B). We use the results from Experiment 2 to specify this distribution. The in-the-moment, contextual informativeness of the utterance is captured in the likelihood term, whose value depends on the rationality parameter $\alpha$. Assumptions about rationality may change depending on context and we therefore used the data from Experiment 1 to specify $\alpha$ (see supplementary information for details about these parameters).

The model generates predictions for situations in which utterance and common ground expectations are jointly manipulated (Fig \ref{fig:fig1}C - see supplementary information for additional details and a worked example of how predictions were generated). In addition to the parameters fit to the data from previous experiments, we include an additional noise parameter to account for responses better explained by a process of random guessing than by pragmatics; we estimate this parameter from the observed data (Experiment 3). Including the noise parameter greatly improved the model fit to the data (see supplementary information for details). We did not pre-register the inclusion of a noise parameter for Experiment 3 but did so for all subsequent experiments.
```{r ex3 GLMM, cache=F}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

In Experiment 3, we combined the procedures of Experiment 1 and 2A or 2B. The test setup was identical to Experiment 1, however, before making a request, the speaker interacted with the objects so that some of them were preferred by or new to them (Fig \ref{fig:fig1}C). We discuss and visualize the results as the proportion with which participants chose the more informative object (i.e., the object that would be the more informative referent when only utterance information is considered). Participants distinguished between congruent and incongruent trials when the speaker remained the same, as evidenced by the fit of a generalized linear mixed effects model (model term: `alignment x speaker`; $\beta$ = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(estimate)`, se = `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(std.error)`, *p* `r lm_ex3_r%>%filter(term == "speakersame_speaker:alignmentincongruent")%>%pull(p.value)`).

```{r fig3, include = T, fig.align = "center", fig.cap = "Results from Experiment 3 and 4 for adults. (A) Data and model predictions by condition for Experiment 3. Transparent dots show data from individual participants, diamonds represent condition means. (B) Correlation between model predictions and data in Experiment 3, (C) between data in Experiment 3 and data for the strong prior manipulation in Experiment 4 (direct replication) and (D) between model predictions and data in Experiment 4. Coefficients and p-values are based on Pearson correlation statistics. Error bars represent 95\\% HDIs." , out.width="400px"}

knitr::include_graphics("./figures/fig3.png")
```

Participants’ average responses were highly correlated with the model’s predictions in each condition (Fig \ref{fig:fig3}B). To test whether participants in fact balanced both information sources, we compared the pragmatics model to two alternative models: the *flat prior model*, which ignores common ground information and the *prior only model*, which ignores utterance information. Model fit was considerably better for the pragmatics model compared to the flat prior model (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information. The estimated proportion of random responses according to the pragmatics model was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This value was substantially lower for the pragmatics model compared to the alternative models (see supplementary information), lending additional support to the conclusion that the pragmatics model better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

### Replication and extension to different levels of common ground information (Experiment 4)

@@

To test if our model makes accurate predictions for different combinations, we first replicated and then extended the results of Experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (strong, medium and weak manipulation) by changing the way the speaker interacted with the objects prior to the request We ran a total of 20 conditions, including a direct replication of Experiment 3 (see Fig \ref{fig:fig3}C).

Model predictions from the pragmatics model were again highly correlated with the average response in each condition (see Fig \ref{fig:fig3}D). We evaluated model fit for the same models as in Experiment 3 and found again that the pragmatics model fit the data much better compared to the flat prior (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the prior only model (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the pragmatics model was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to Experiment 3 and again lower compared to the alternative models (see supplementary information).

## Do children integrate contextual information?

@@

### Inferences based on utterance and common ground information (Experiment 5 and 6)

The study for children followed the same general pattern as the one for adults. We generated model predictions for how information should be integrated by first measuring children’s ability to use utterance (informativeness) and common ground (preference) information in isolation when making pragmatic inferences. We then adapted our model to study developmental change: We sampled children continuously between 3.0 and 5.0 years of age – a time in which children have been found to make the kind of pragmatic inferences we studied here [@bohn2019pervasive; frank2014inferring] - and generated model predictions for the average developmental trajectory in each condition\footnote{For Experiment 5 and 6, we also tested two-year-olds but did not find sufficient evidence that they use utterance and/or common ground information in the tasks we used to justify investigating their ability to integrate the two.}.

@@

All experiments were framed as games in which participants would learn words from animals. They were implemented in HTML/JavaScript as a website. Adults were directed to the website via MTurk and responded by clicking objects. Children were guided through the game by an experimenter and responded by touching objects on the screen of an iPad tablet [@frank2016using]. For each animal character, we recorded a set of utterances (one native English speaker per animal) that were used to provide information and make requests. All experiments started with an introduction to the animals and two training trials in which familiar objects were requested (car and ball). Subsequent test trials in each condition were presented in a random order.
